---
title: "simulate_outcome"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simulate_outcome}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
# knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(rdborrow)
set.seed(2023)
```

In this document we discuss how to simulate outcomes for experimentation. We will discuss two parts: 

1. Simulate outcomes for primary analysis; 

2. Simulate outcomes under OLE settings;




## 1. Simulate outcomes for primary analysis
We use the following generic setup to facilitate the discussion: 

1. Number of patients: $N = 300$, with $N_1 = 200$ in internal study and $N_0 = 100$ in external study;
2. Covariates: 2 categorical variables ($\text{p_cat} = 2$) and 2 continuous variables ($\text{p_cont} = 2$), from a discretized multivarite gaussian model;
3. Treatment allocation: for internal group $S = 1$ , treatment $A$ is randomized following $\text{Bernoulli}(2/3)$; for external group $S = 0$, all units are control units, i.e. with $A = 0$;
4. Followup time: two follow up time points;
5. Potential outcomes models: at each time point, the outcomes generated from linear models. 


### Simulate $X$ for $S = 0$ and $S = 1$

The first step is to simulate covariate $X$ for both the internal group $S = 0$ and the external group $S = 1$. Two groups are allowed to have different sample sizes, but their columns should represent the same set of variables. 

```{r}
# separate is better: could be another S
# TODO: checking the same number of dimension

# S = 1
X_int = simulate_X_dct_mvnorm(n = 200,   # number of units
                              p = 4,    # dimension of variables
                              mu = rep(1, 4),  # population mean of multivariate normal
                              sig = diag(4),   # population covariance matrix of mv normal
                              cat_cols = c(1, 2), # which columns are categorical
                              cat_prob = list(c(0.3, 0.7), c(0.4, 0.6)) # the pdf of categorical variables 
                              )

# S = 0
X_ext = simulate_X_dct_mvnorm(n = 100,   # number of units
                              p = 4,    # dimension of variables
                              mu = rep(0, 4),  # population mean of multivariate normal
                              sig = diag(4),   # population covariance matrix of mv normal
                              cat_cols = c(1, 2), # which columns are categorical
                              cat_prob = list(c(0.3, 0.7), c(0.4, 0.6)) # the pdf of categorical variables 
                              )
```

Display the first 10 rows of the covariate matrices:
```{r}
head(X_int, 10)
head(X_ext, 10)
```

### Simulate trial
```{r}
# Specify outcome models
model_form_x_t1 = "x1*1 + x2*1 + x3*1 + x4*1"
model_form_x_t2 = "x1*1 + x2*(-1) + x3*(-1) + x4*1"

outcome_model_specs = list(
  list(effect = 3, model_form_x = model_form_x_t1, noise_mean = 0, noise_sd = 0.5),  # model form for the first time point, given by model_form1
  list(effect = 0, model_form_x = model_form_x_t2, noise_mean = 0, noise_sd = 0.5)  # model form for the second time point, given by model_form2
)

Data = simulate_trial(X_int, X_ext, prob_treated = 2/3, OLE_flag = F, T_cross = 0, outcome_model_specs)
```

Display the first 10 rows of the dataset:
```{r}
head(Data, 10)
```




## 2. Simulate outcomes for OLE study

For OLE study, while most of the components remain unchanged compared to primary analysis, there are two crucial modifications:

(1) For treatment allocation, besides the allocation indicator for the first phase, we also need to add in one indicator for the second phase;
(2) For outcome model, the model forms should be formulated in terms of the treatment allocation vector specific to each phase of the design.

Here, besides the aforementioned setup, we further consider two more additional followup time points. At these two later time points, all the participants in the interal group are switched to treated group $A = 1$. The outcome models are still linear but with time-varying parameters.


### Simulate outcomes after OLE
```{r}
# Specify outcome models after OLE
# clarify certain format of the dataset
model_form_x_t1 = "x1*1 + x2*1 + x3*1 + x4*1"
model_form_x_t2 = "x1*1 + x2*(-1) + x3*(-1) + x4*1"
model_form_x_t3 = "x1*1 + x2*1 + x3*1 + x4*1"
model_form_x_t4 = "x1*1 + x2*(-1) + x3*(-1) + x4*1"

outcome_model_specs = list(
  list(effect = 3, model_form_x = model_form_x_t1, noise_mean = 0, noise_sd = 0.5),  # model form for the first time point, given by model_form1
  list(effect = 0, model_form_x = model_form_x_t2, noise_mean = 0, noise_sd = 0.5),  # model form for the second time point, given by model_form2
  list(effect = 5, model_form_x = model_form_x_t3, noise_mean = 0, noise_sd = 0.5),  # model form for the third time point, given by model_form3
  list(effect = 5, model_form_x = model_form_x_t4, noise_mean = 0, noise_sd = 0.5)   # model form for the fourth time point, given by model_form4
)


Data = simulate_trial(X_int,   # simulated covariates for internal study
                      X_ext,   # simulated covariates for external study
                      prob_treated = 2/3, # treatment probability for two periods
                      OLE_flag = T,
                      T_cross = 2,  # follow up time within each phases
                      outcome_model_specs)  # outcome models

```


Display the first 10 rows of the dataset:
```{r}
head(Data, 10)
```

putting some instruction here...

