---
title: "simulate_covariates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simulate_covariates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
# knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(rdborrow)
set.seed(2023)
```


# Simulating Covariates

## Three ways for generating covariates:

**1. Simulating data from discretized multivariate normal vectors**

Suppose we want to generate a data that has `p` variables, with `p_cat` numbers of categorical variables and `p_cont = p - p_cat` continuous variables. 

- Generate a $p$-dimensional multivariate gaussian vector:
$$
X = (X_1,\dots,X_p)^\top \sim N_p(\mu, \Sigma)
$$
- Pick \texttt{p_cat} columns and truncate them using quantiles of normal distribution. For example, to discretize $X_1$ into a binary variable with $P(X_1 = 0) = 0.3, P(X_1 = 1) = 0.7$, we choose the $0.7$ upper quantile $q_{0.7}$ and truncate $X_1$:
$$
X_1 = 1(X_1 \ge q_{0.7}).
$$

**2. Simulating data from mixture models**

Another way to simulate $X$ is from gaussian mixture models. Such models have great interpretability. Some continuous variables are distributed according to levels of categorical variables. For example, heights may depend on gender; grading results may depend on subjects; etc. 

To simulate $X$ from mixture gaussian distributions, we proceed as follows:
- Specify joint distribution of `p_cat` categorical variables:
$$
P(X_1 = x_1, \dots, X_{\text{p_cat}} = x_{\text{p_cat}}) = p
$$
- For each level combination of categorical variables, generate a joint normal distribution with given mean $\mu$ and covariance $\Sigma$:
$$
P(X_{\text{p_cat} + 1}, \dots, X_p \mid X_1 = x_1, \dots, X_{\text{p_cat}} = x_{\text{p_cat}}) = N_{\text{p_cont}}(\mu, \Sigma)
$$

**3. Simulating data from copula with specified marginal distributions**

Copula provides a powerful way to couple a given set of marginal distributions. In implementation, we follow the following procedure to generate data from a given copula:

- Specify a copula model, say normal or t copula (elliptical copula), Clayton copula (Archimedean copula), Gumbel copula (extreme value copula), etc.  

- Specify the marginal distributions: say normal, t, binomial, etc.

TODO: in certain situation use which

## Implementing three methods under a specific setup
In the following sections, we consider a running setup to simulate $X$ with:

- $n = 100$ patients;
- $\text{p_cat} = 2$ categorical covariates; one has two levels and the other has three levels;
- $\text{p_con} = 2$ continuous covariates.


### 1. Simulating data from discretized multivariate normal vectors

We can simulate $X$ from discretized normal distribution. This example can be implemented by the following code chunk:
```{r echo = TRUE}
# testing 
# put some seed
# sanity check
test1 = simulate_X_dct_mvnorm(n = 100,   # number of units
                              p = 4,    # dimension of variables
                              mu = rep(0, 4),  # population mean of multivariate normal
                              sig = diag(4),   # population covariance matrix of mv normal
                              cat_cols = c(1, 2), # which columns are categorical
                              cat_prob = list(c(0.3, 0.7), c(0.1, 0.3, 0.6)) # the pdf of categorical variables 
                              )
head(test1, 10)
```



    

### 2. Simulating data from mixture models

Simulate X from a gaussian mixture model:
```{r}
# test mixture of continuous and categorical
X = simulate_X_mixture(n = 100, p_cat = 2, p_cont = 2, 
                   cat_level_list = list(x1 = c(0,1), x2 = c(0,1,2)),
                   cat_comb_prob = c(0.1, 0.1, 0.3, 0.3, 0.1, 0.1),
                   cont_para_list = list(list(mean = c(0, 0), sigma = diag(2)), 
                                         list(mean = c(0, 1), sigma = diag(2)),
                                         list(mean = c(0, 2), sigma = diag(2)),
                                         list(mean = c(1, 0), sigma = diag(2)),
                                         list(mean = c(1, 1), sigma = diag(2)),
                                         list(mean = c(1, 2), sigma = diag(2))
                                         )
                   )
head(X, 10)
```


    
    
    
### 3. Simulating data from copula with specified marginal distributions

Explain the parameters

```{r}
# generate a 4-dimensional normal copula with an AR1 covariance structure
normal <- copula::normalCopula(param = c(0.8), dim = 4, dispstr = "ar1")

# simulate 100 sample
X <- simulate_X_copula(n = 100, 
                       p = 4, 
                       cp = normal,  # copula
                       margins = c("binom", "binom", "norm", "norm"),  # specify marginal distributions
                       paramMargins = list(list(size = 2, prob = 0.5), # specify parameters for marginals
                                  list(size = 3, prob = 0.5),
                                  list(mean = 0, sd = 1),
                                  list(mean = 0, sd = 1))
                       )
head(X, 10)
```


## Comparing Pros and Cons of the simulating methods

| Methods              | Pros                                                                                                                                                                                                                                           | Cons                                                                                                                       |
|----------------------|------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|----------------------------------------------------------------------------------------------------------------------------|
| Discretized Gaussian | - easy to implement; <br>- easy to specify columns and pdf of categorical variables;<br>- can specify correlation structure                                                                                                                    | - continuous variables are limited to normal;<br>- hard to directly quantify the correlation between categorical varialbes |
| Gaussian Mixture     | - highly interpretable: levels of categorical variables form clusters, each cluster with one distribution (pattern)<br>- very flexible for specifying correlation structure of continuous variables and joint pattern of categorical variables | - might have many parameters to specify with many covariates                                                               |
| Copula               | - can accommodate many different marginal distributions, such as t, normal, gamma, etc.<br>- instead of specifying Pearson correlation, can specify Spearman correlation                                                                       | - Spearman correlation is not an accurate characterization for correlation among categorical variables                     |
    

## Additional examples

### Using copula methods to generate non-normal marginals

We can generate a multivariate distribution with non-normal marginals using copula. In the following example, we generate data that couples two normal distributions, one t distribution and one poisson distribution. 

```{r}
# test simulation from copula
normal <- copula::normalCopula(param = c(0.8), dim = 4, dispstr = "ar1")
X <- simulate_X_copula(100, 4, normal, 
                       margins = c("norm", "t", "norm", "pois"), 
                       paramMargins = list(list(mean = 2, sd=3),
                                  list(df = 2),
                                  list(mean = 0, sd = 1),
                                  list(lambda = 5))
                       )
head(X, 10)
```
