---
title: "Primary Analysis Workflow"
author: "Lei"
date: "2024-02-19"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(rdborrow)
```



## Primary analysis
### 1 load and visualize data
```{r}
# load the simulated dataset
head(SyntheticData)
```


### 2 Estimation and inference:

#### 2.1 Inverse probability weighting (IPW)
1) IPW with zero weight (wt = 0):

```{r}
# test: within trial
## Data argument + column names (coxph, glm)
method_weighting_obj = setup_method_weighting(
  method_name = "IPW",
  optimal_weight_flag = F, 
  wt = 0,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5")

analysis_primary_obj = setup_analysis_primary(
  data = SyntheticData,
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  method_weighting_obj = method_weighting_obj)

res = run_analysis(analysis_primary_obj)
res
```

2) IPW with data-adaptive weight:
```{r}
# test: within trial
method_weighting_obj = setup_method_weighting(
  method_name = "IPW",
  optimal_weight_flag = T,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5")

analysis_primary_obj = setup_analysis_primary(
  data = SyntheticData,
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  method_weighting_obj = method_weighting_obj)

res = run_analysis(analysis_primary_obj)

res$borrow_weight
res$results
```

#### 2.2 Augmented inverse probability weighting (AIPW)

The second approach is AIPW, which also accommodates two external borrowing strategies.

1) AIPW with zero weight (wt = 0):
```{r}
# test: AIPW with 0 weight, should be same as IPW with 0 weight
method_weighting_obj = setup_method_weighting(
  method_name = "AIPW",
  optimal_weight_flag = F, 
  wt = 0,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
  model_form_mu0_ext = c("y1 ~ x1 + x2 + x3 + x4 + x5",
                         "y2 ~ x1 + x2 + x3 + x4 + x5"))

analysis_primary_obj = setup_analysis_primary(
  data = SyntheticData,
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  method_weighting_obj = method_weighting_obj)

res = run_analysis(analysis_primary_obj)
```



2) AIPW with data adaptive weight:
```{r}
# test: AIPW with given weight
# bootstrap as part of the method and analysis
method_weighting_obj = setup_method_weighting(
  method_name = "AIPW",
  optimal_weight_flag = T,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
  model_form_mu0_ext = c("y1 ~ x1 + x2 + x3 + x4 + x5",
                         "y2 ~ x1 + x2 + x3 + x4 + x5"))

analysis_primary_obj = setup_analysis_primary(
  data = SyntheticData,
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  method_weighting_obj = method_weighting_obj)

res = run_analysis(analysis_primary_obj)
```



### 3 Bootstrap inference

In this section we present Bootstrap inference results. We report Bootstrap confidence intervals with adjusted quantile ranges. 

1) IPW with bootstrap CI
```{r}
# test: within trial
## Data argument + column names (coxph, glm)
## having a bootstrap_flag outside the class
bootstrap_obj = setup_bootstrap(
  replicates = 2e3,
  bootstrap_CI_type = "perc"
)

method_weighting_obj = setup_method_weighting(
  method_name = "IPW",
  optimal_weight_flag = T, 
  bootstrap_flag = T,
  bootstrap_obj = bootstrap_obj,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5")

analysis_primary_obj = setup_analysis_primary(
  data = SyntheticData,
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  method_weighting_obj = method_weighting_obj)

res = run_analysis(analysis_primary_obj)
res

```

2) AIPW with bootstrap CI

```{r}
# test: with optimal weight
## Data argument + column names (coxph, glm)
bootstrap_obj = setup_bootstrap(
  replicates = 2e3,
  bootstrap_CI_type = "perc"
)

method_weighting_obj = setup_method_weighting(
  method_name = "AIPW",
  optimal_weight_flag = T, 
  wt = NaN,
  bootstrap_flag = T,
  bootstrap_obj = bootstrap_obj,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
  model_form_mu0_ext = c("y1 ~ x1 + x2 + x3 + x4 + x5",
                         "y2 ~ x1 + x2 + x3 + x4 + x5"))

analysis_primary_obj = setup_analysis_primary(
  data = SyntheticData,
  trial_status = "S", 
  treatment = "A", 
  outcome = c("y1", "y2"), 
  covariates = c("x1", "x2", "x3", "x4", "x5"), 
  method_weighting_obj = method_weighting_obj)


res = run_analysis(analysis_primary_obj)

## best to just do the last one, but also have options
## time dependent way of effects and modeling 
```

### 4 Notes

1. When there are missing values in the data, the suggestion we have for now is to preprocess the dataset (such as deletion, imputing, etc.) to obtain a dataset without missingness, then apply the package. For general methodology development regarding missing values, we save it for future research work.  