---
title: "replication"
author: "Lei"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(rdborrow)
```

This document summarizes the replication details of the numerical experiments in the submitted manuscripts. 

## Primary analysis (first subsection in Section 4.1)
### 1 load and visualize data
```{r}
# Initialize an empty data list
head(SyntheticData)
```


### 2 Estimation and inference:

#### 2.1 Inverse probability weighting (IPW)
1) IPW with zero weight (wt = 0):

```{r}
# test: within trial
## Data argument + column names (coxph, glm)
method_weighting_obj = setup_method_weighting(
  method_name = "IPW",
  optimal_weight_flag = F, 
  wt = 0,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5")

analysis_primary_obj = setup_analysis_primary(
  data = SyntheticData,
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  method_weighting_obj = method_weighting_obj)

res = run_analysis(analysis_primary_obj)
res
```

2) IPW with data-adaptive weight:
```{r}
# test: within trial
method_weighting_obj = setup_method_weighting(
  method_name = "IPW",
  optimal_weight_flag = T,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5")

analysis_primary_obj = setup_analysis_primary(
  data = SyntheticData,
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  method_weighting_obj = method_weighting_obj)

res = run_analysis(analysis_primary_obj)

res$borrow_weight
res$results
```

#### 2.2 Augmented inverse probability weighting (AIPW)

The second approach is AIPW, which also accommodates two external borrowing strategies.

1) AIPW with zero weight (wt = 0):
```{r}
# test: AIPW with 0 weight, should be same as IPW with 0 weight
method_weighting_obj = setup_method_weighting(
  method_name = "AIPW",
  optimal_weight_flag = F, 
  wt = 0,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
  model_form_mu0_ext = c("y1 ~ x1 + x2 + x3 + x4 + x5",
                         "y2 ~ x1 + x2 + x3 + x4 + x5"))

analysis_primary_obj = setup_analysis_primary(
  data = SyntheticData,
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  method_weighting_obj = method_weighting_obj)

res = run_analysis(analysis_primary_obj)
```



2) AIPW with data adaptive weight:
```{r}
# test: AIPW with given weight
# bootstrap as part of the method and analysis
method_weighting_obj = setup_method_weighting(
  method_name = "AIPW",
  optimal_weight_flag = T,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
  model_form_mu0_ext = c("y1 ~ x1 + x2 + x3 + x4 + x5",
                         "y2 ~ x1 + x2 + x3 + x4 + x5"))

analysis_primary_obj = setup_analysis_primary(
  data = SyntheticData,
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  method_weighting_obj = method_weighting_obj)

res = run_analysis(analysis_primary_obj)
```



### 3 Bootstrap inference

In this section we present Bootstrap inference results. We report Bootstrap confidence intervals with adjusted quantile ranges. 

1) IPW with bootstrap CI
```{r}
# test: within trial
## Data argument + column names (coxph, glm)
## having a bootstrap_flag outside the class
bootstrap_obj = setup_bootstrap(
  replicates = 2e3,
  bootstrap_CI_type = "perc"
)

method_weighting_obj = setup_method_weighting(
  method_name = "IPW",
  optimal_weight_flag = T, 
  bootstrap_flag = T,
  bootstrap_obj = bootstrap_obj,
  wt = 0,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5")

analysis_primary_obj = setup_analysis_primary(
  data = SyntheticData,
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  method_weighting_obj = method_weighting_obj)

res = run_analysis(analysis_primary_obj)
res

```

2) AIPW with bootstrap CI

```{r}
# test: with optimal weight
## Data argument + column names (coxph, glm)
bootstrap_obj = setup_bootstrap(
  replicates = 2e3,
  bootstrap_CI_type = "perc"
)

method_weighting_obj = setup_method_weighting(
  method_name = "AIPW",
  optimal_weight_flag = T, 
  wt = 0,
  bootstrap_flag = T,
  bootstrap_obj = bootstrap_obj,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
  model_form_mu0_ext = c("y1 ~ x1 + x2 + x3 + x4 + x5",
                         "y2 ~ x1 + x2 + x3 + x4 + x5"))

analysis_primary_obj = setup_analysis_primary(
  data = SyntheticData,
  trial_status = "S", 
  treatment = "A", 
  outcome = c("y1", "y2"), 
  covariates = c("x1", "x2", "x3", "x4", "x5"), 
  method_weighting_obj = method_weighting_obj)


res = run_analysis(analysis_primary_obj)

## best to just do the last one, but also have options
## time dependent way of effects and modeling 
```





## OLE phase analysis (second subsection in Section 4.1)

### 1 DID methods

#### 1.1 IPW
```{r message=FALSE, warning=FALSE}
bootstrap_obj = setup_bootstrap(
  replicates = 2e3, 
  bootstrap_CI_type = "perc" 
)

method_DID_obj = setup_method_DID(
  method_name = "IPW",
  bootstrap_flag = T,
  bootstrap_obj = bootstrap_obj,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
  model_form_piA = "A ~ x1 + x2 + x3 + x4 + x5")

analysis_OLE_obj = setup_analysis_OLE( 
  data = SyntheticData, 
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2", "y3", "y4"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  T_cross = 2, 
  method_OLE_obj = method_DID_obj)

res = run_analysis(analysis_OLE_obj)
```

#### 1.2 AIPW
```{r message=FALSE, warning=FALSE}
bootstrap_obj = setup_bootstrap(
  replicates = 2e3, 
  bootstrap_CI_type = "perc" 
)

model_form_mu = c("y1 ~ x1 + x2 + x3 + x4 + x5", 
                  "y2 ~ x1 + x2 + x3 + x4 + x5",
                  "y3 ~ x1 + x2 + x3 + x4 + x5",
                  "y4 ~ x1 + x2 + x3 + x4 + x5")

method_DID_obj = setup_method_DID(
  method_name = "AIPW",
  bootstrap_flag = T,
  bootstrap_obj = bootstrap_obj,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
  model_form_piA = "A ~ x1 + x2 + x3 + x4 + x5",
  model_form_mu0_ext = model_form_mu)

analysis_OLE_obj = setup_analysis_OLE( 
  data = SyntheticData, 
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2", "y3", "y4"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  T_cross = 2,  
  method_OLE_obj = method_DID_obj) 

res = run_analysis(analysis_OLE_obj) 
```


#### 1.3 OR
```{r message=FALSE, warning=FALSE}
bootstrap_obj = setup_bootstrap(
  replicates = 2e3, 
  bootstrap_CI_type = "perc" 
)

model_form_mu = c("y1 ~ x1 + x2 + x3 + x4 + x5", 
                  "y2 ~ x1 + x2 + x3 + x4 + x5",
                  "y3 ~ x1 + x2 + x3 + x4 + x5",
                  "y4 ~ x1 + x2 + x3 + x4 + x5")

method_DID_obj = setup_method_DID(
  method_name = "OR",
  bootstrap_flag = T,
  bootstrap_obj = bootstrap_obj,
  model_form_mu0_ext = model_form_mu, 
  model_form_mu0_rct = model_form_mu,
  model_form_mu1_rct = model_form_mu)

analysis_OLE_obj = setup_analysis_OLE(
  data = SyntheticData,
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2", "y3", "y4"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"),
  T_cross = 2,
  method_OLE_obj = method_DID_obj)

res = run_analysis(analysis_OLE_obj)
res
```


### 2 SCM method: with parallel computing
```{r}
bootstrap_obj = setup_bootstrap(
  replicates = 200, 
  bootstrap_CI_type = "perc" 
)

method_SCM_obj = setup_method_SCM(
  method_name = "SCM", 
  bootstrap_flag = T, 
  bootstrap_obj = bootstrap_obj, 
  lambda.min = 0, 
  lambda.max = 1e-3, 
  nlambda = 10, 
  parallel = "multicore", 
  ncpus = 4) 

analysis_OLE_obj = setup_analysis_OLE(
  data = SyntheticData, 
  trial_status_col_name = "S", 
  treatment_col_name = "A", 
  outcome_col_name = c("y1", "y2", "y3", "y4"), 
  covariates_col_name = c("x1", "x2", "x3", "x4", "x5"), 
  T_cross = 2, 
  method_OLE_obj = method_SCM_obj) 

run_analysis(analysis_OLE_obj) 
```


## Simulation for primary analysis (first subsection in Section 4.2)
### 1 Simulate a list of datasets for primary analysis

#### 1.1 Create basic setup
```{r}
# Initialize an empty data list
set.seed(2023)

data_matrix_list_null = list()
data_matrix_list_alt = list()
ntrial = 500

# Specify the significance level alpha 
alpha = 0.05

# Specify the true effect size at the end of the study
true_effect = 0
alt_effect = 2.0 # tune this

# Specify column names
covariates_col_name = c("x1", "x2", "x3", "x4", "x5")
outcome_col_name = c("y1", "y2")
treatment_col_name = "A"
trial_status_col_name = "S"
```


#### 1.2 Generate simulation data list for null hypothesis
```{r}
# Sequentially adding in datasets
for (trial_iter in 1:ntrial){
  # simulate 300 sample
  normal <- copula::normalCopula(param = c(0.8), dim = 4, dispstr = "ar1")
  
  #========== generate internal covariates =============
  X_int <- simulate_X_copula(n = 200, 
                         p = 4, 
                         cp = normal,  # copula
                         margins = c("binom", "binom", "binom", "exp"),  # specify marginal distributions
                         paramMargins = list(list(size = 1, prob = 0.7), # specify parameters for marginals
                                             list(size = 1, prob = 0.9), 
                                             list(size = 1, prob = 0.3), 
                                             list(rate = 1/10)) 
  ) 
  
  X_int$x4 = round(X_int$x4) + 1
  X_int$x5 = 30 + 10 * X_int$x1 + (7) * X_int$x2 + (-6) * X_int$x3 + (-0.5) * X_int$x4 + rnorm(200, mean = 0, sd = 10)
  
  varnames = c("1", paste0("x", 1:5))
  
  #============ generate external covariates ==============
  X_ext <- simulate_X_copula(n = 100, 
                             p = 4, 
                             cp = normal,  # copula
                             margins = c("binom", "binom", "binom", "exp"),  # specify marginal distributions
                             paramMargins = list(list(size = 1, prob = 0.7), # specify parameters for marginals
                                                 list(size = 1, prob = 0.9), 
                                                 list(size = 1, prob = 0.3), 
                                                 list(rate = 1/10)) 
  ) 
  
  X_ext$x4 = round(X_ext$x4) + 1
  X_ext$x5 = 50 + 10 * X_ext$x1 + (2) * X_ext$x2 + (-1) * X_ext$x3 + (-0.3) * X_ext$x4 + rnorm(100, mean = 0, sd = 10)
  
  varnames = c("1", paste0("x", 1:5))
  
  #============ Specify outcome models ==============
  model_form_x_t1 = setNames(c(10.0, 0.05, -1.5, -1.0, -0.2, -0.1), varnames) # 1.5*A, sigma = 4.0
  model_form_x_t2 = setNames(c(6.0, 0.5, -0.5, -1.0, -0.3, -0.06), varnames) # 1.8*A, sigma = 4.0
  # model_form_x_t3 = setNames(c(5.0, 1.9, 1.4, -1.3, -0.4, -0.15), varnames) # 1.6*A, sigma = 4.0
  # model_form_x_t4 = setNames(c(1.2, 1.0, 2.0, -0.5, -0.4, -0.10), varnames) # 2.5*A, sigma = 5.0
  
  outcome_model_specs = list(
    list(effect = 0, model_form_x = model_form_x_t1, # from data: true_effect = 1.5
         noise_mean = 0, noise_sd = 4),  # model form for the first time point, given by model_form1
    list(effect = true_effect, model_form_x = model_form_x_t2, # from data: true_effect = 1.8
         noise_mean = 0, noise_sd = 4)  # model form for the second time point, given by model_form2
  )
  
  #=========== generate trial data ============
  Data = simulate_trial(X_int, 
                        X_ext, 
                        num_treated = 100, 
                        OLE_flag = F, 
                        T_cross = 2,  
                        outcome_model_specs)
  
  data_matrix_list_null[[trial_iter]] = Data
}



print(data_matrix_list_null[[1]])

```



#### 1.3 Generate simulation data list for alternative hypothesis
```{r}
# Sequentially adding in datasets
for (trial_iter in 1:ntrial){
  # simulate 300 sample
  normal <- copula::normalCopula(param = c(0.8), dim = 4, dispstr = "ar1")
  
  #========== generate internal covariates =============
  X_int <- simulate_X_copula(n = 200, 
                         p = 4, 
                         cp = normal,  # copula
                         margins = c("binom", "binom", "binom", "exp"),  # specify marginal distributions
                         paramMargins = list(list(size = 1, prob = 0.7), # specify parameters for marginals
                                             list(size = 1, prob = 0.9), 
                                             list(size = 1, prob = 0.3), 
                                             list(rate = 1/10)) 
  ) 
  
  X_int$x4 = round(X_int$x4) + 1
  X_int$x5 = 30 + 10 * X_int$x1 + (7) * X_int$x2 + (-6) * X_int$x3 + (-0.5) * X_int$x4 + rnorm(200, mean = 0, sd = 10)
  
  varnames = c("1", paste0("x", 1:5))
  
  #============ generate external covariates ==============
  X_ext <- simulate_X_copula(n = 100, 
                             p = 4, 
                             cp = normal,  # copula
                             margins = c("binom", "binom", "binom", "exp"),  # specify marginal distributions
                             paramMargins = list(list(size = 1, prob = 0.7), # specify parameters for marginals
                                                 list(size = 1, prob = 0.9), 
                                                 list(size = 1, prob = 0.3), 
                                                 list(rate = 1/10)) 
  ) 
  
  X_ext$x4 = round(X_ext$x4) + 1
  X_ext$x5 = 50 + 10 * X_ext$x1 + (2) * X_ext$x2 + (-1) * X_ext$x3 + (-0.3) * X_ext$x4 + rnorm(100, mean = 0, sd = 10)
  
  varnames = c("1", paste0("x", 1:5))
  
  #============ Specify outcome models ==============
  model_form_x_t1 = setNames(c(10.0, 0.05, -1.5, -1.0, -0.2, -0.1), varnames) # 1.5*A, sigma = 4.0
  model_form_x_t2 = setNames(c(6.0, 0.5, -0.5, -1.0, -0.3, -0.06), varnames) # 1.8*A, sigma = 4.0
  # model_form_x_t3 = setNames(c(5.0, 1.9, 1.4, -1.3, -0.4, -0.15), varnames) # 1.6*A, sigma = 4.0
  # model_form_x_t4 = setNames(c(1.2, 1.0, 2.0, -0.5, -0.4, -0.10), varnames) # 2.5*A, sigma = 5.0
  
  outcome_model_specs = list(
    list(effect = 0, model_form_x = model_form_x_t1, # from data: true_effect = 1.5
         noise_mean = 0, noise_sd = 4),  # model form for the first time point, given by model_form1
    list(effect = alt_effect, model_form_x = model_form_x_t2, # from data: true_effect = 1.8
         noise_mean = 0, noise_sd = 4)  # model form for the second time point, given by model_form2
  )
  
  #=========== generate trial data ============
  Data = simulate_trial(X_int, 
                        X_ext, 
                        num_treated = 100, 
                        OLE_flag = F, 
                        T_cross = 2,  
                        outcome_model_specs)
  
  data_matrix_list_alt[[trial_iter]] = Data
}

data_matrix_list_alt[[1]]
```


### 2 Parametric inference:
```{r}
# Specify a list of methods to be tested
## IPW - optimal weight
method_IPW_optimal_weight = setup_method_weighting(
  method_name = "IPW",
  optimal_weight_flag = T,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5")

## AIPW - optimal weight
method_AIPW_optimal_weight = setup_method_weighting(
  method_name = "AIPW",
  optimal_weight_flag = T, 
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
  model_form_mu0_ext = c("y1 ~ x1 + x2 + x3 + x4 + x5",
                         "y2 ~ x1 + x2 + x3 + x4 + x5"))

## IPW - zero weight
method_IPW_zero_weight = setup_method_weighting(
  method_name = "IPW",
  wt = 0, 
  optimal_weight_flag = F,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5")

## AIPW - zero weight
method_AIPW_zero_weight = setup_method_weighting(
  method_name = "AIPW",
  wt = 0,
  optimal_weight_flag = F, 
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
  model_form_mu0_ext = c("y1 ~ x1 + x2 + x3 + x4 + x5",
                         "y2 ~ x1 + x2 + x3 + x4 + x5")) 

## method list
method_obj_list = list(method_IPW_optimal_weight, 
                       method_AIPW_optimal_weight,
                       method_IPW_zero_weight,
                       method_AIPW_zero_weight)


```


```{r}
# create a simulation object for primary analysis
simulation_primary_obj = setup_simulation_primary(
  data_matrix_list_null = data_matrix_list_null,  # two scenarios
  data_matrix_list_alt = data_matrix_list_alt,
  trial_status_col_name = trial_status_col_name, 
  treatment_col_name = treatment_col_name, 
  outcome_col_name = outcome_col_name, 
  covariates_col_name = covariates_col_name, 
  method_obj_list = method_obj_list, 
  true_effect = true_effect,  
  alt_effect = alt_effect,
  alpha = alpha,
  method_description = c("IPW, optimal weight", 
                         "AIPW, optimal weight", 
                         "IPW, zero weight",
                         "AIPW, zero weight"))
```

```{r}
simulation_report = run_simulation(simulation_primary_obj, quiet = F)
```

```{r}
simulation_report # Type I error and Power
```



### 3 Bootstrap inference:
```{r}
# Specify a list of methods to be tested
bootstrap_obj = setup_bootstrap(
  replicates = 5e2,
  bootstrap_CI_type = "bca"
)

## IPW - optimal weight
method_IPW_optimal_weight = setup_method_weighting(method_name = "IPW",
                                                   optimal_weight_flag = T,
                                                   bootstrap_flag = T,
                                                   bootstrap_obj = bootstrap_obj,
                                                   model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5")

## AIPW - optimal weight
method_AIPW_optimal_weight = setup_method_weighting(method_name = "AIPW",
                                                    optimal_weight_flag = T, 
                                                    bootstrap_flag = T,
                                                    bootstrap_obj = bootstrap_obj,
                                                    model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
                                                    model_form_mu0_ext = c("y1 ~ x1 + x2 + x3 + x4 + x5",
                                                                           "y2 ~ x1 + x2 + x3 + x4 + x5 + y1"))

## IPW - zero weight
method_IPW_zero_weight = setup_method_weighting(method_name = "IPW",
                                                wt = 0, 
                                                optimal_weight_flag = F,
                                                bootstrap_flag = T,
                                                bootstrap_obj = bootstrap_obj,
                                                model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5")

## AIPW - zero weight
method_AIPW_zero_weight = setup_method_weighting(
  method_name = "AIPW",
  wt = 0,
  optimal_weight_flag = F, 
  bootstrap_flag = T,
  bootstrap_obj = bootstrap_obj,
  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
  model_form_mu0_ext = c("y1 ~ x1 + x2 + x3 + x4 + x5",
                         "y2 ~ x1 + x2 + x3 + x4 + x5 + y1")) 

## method list
method_obj_list = list(method_IPW_optimal_weight, 
                       method_AIPW_optimal_weight,
                       method_IPW_zero_weight,
                       method_AIPW_zero_weight)

```


```{r}
# create a simulation object for primary analysis
simulation_primary_obj = setup_simulation_primary(
  data_matrix_list_null = data_matrix_list_null, 
  data_matrix_list_alt = data_matrix_list_alt,
  trial_status_col_name = trial_status_col_name, 
  treatment_col_name = treatment_col_name, 
  outcome_col_name = outcome_col_name, 
  covariates_col_name = covariates_col_name, 
  method_obj_list = method_obj_list, 
  true_effect = true_effect, 
  alt_effect = alt_effect, 
  alpha = alpha,
  method_description = c("IPW, optimal weight, bootstrap", 
                         "AIPW, optimal weight, bootstrap", 
                         "IPW, zero weight, bootstrap",
                         "AIPW, zero weight, bootstrap"))
```

```{r}
simulation_report_bs = run_simulation(simulation_primary_obj, quiet = FALSE)
```

```{r}
simulation_report_bs
```




## Simulation for the OLE phase  (second subsection in Section 4.2)

### 1. Simulate a dataset for OLE study
```{r}
# Initialize an empty data list
set.seed(2023)

data_matrix_list = list()
ntrial = 20

# Specify the significance level alpha 
alpha = 0.05

# Specify the true effect size at the end of the OLE study
true_effect_long = 0

# Specify column names
covariates_col_name = c("x1", "x2", "x3", "x4", "x5")
outcome_col_name = c("y1", "y2", "y3", "y4")
treatment_col_name = "A"
trial_status_col_name = "S"
```



```{r}
# Sequentially adding in datasets
for (trial_iter in 1:ntrial){
  # simulate 300 sample
  normal <- copula::normalCopula(param = c(0.8), dim = 4, dispstr = "ar1")
  
  #========== generate internal covariates =============
  X_int <- simulate_X_copula(n = 200, 
                             p = 4, 
                             cp = normal,  # copula
                             margins = c("binom", "binom", "binom", "exp"),  # specify marginal distributions
                             paramMargins = list(list(size = 1, prob = 0.7), # specify parameters for marginals
                                                 list(size = 1, prob = 0.9), 
                                                 list(size = 1, prob = 0.3), 
                                                 list(rate = 1/10)) 
  ) 
  
  X_int$x4 = round(X_int$x4) + 1
  X_int$x5 = 30 + 10 * X_int$x1 + (7) * X_int$x2 + (-6) * X_int$x3 + (-0.5) * X_int$x4 + rnorm(200, mean = 0, sd = 10)
  
  varnames = c("1", paste0("x", 1:5))
  
  #============ generate external covariates ==============
  X_ext <- simulate_X_copula(n = 100, 
                             p = 4, 
                             cp = normal,  # copula
                             margins = c("binom", "binom", "binom", "exp"),  # specify marginal distributions
                             paramMargins = list(list(size = 1, prob = 0.7), # specify parameters for marginals
                                                 list(size = 1, prob = 0.9), 
                                                 list(size = 1, prob = 0.3), 
                                                 list(rate = 1/10)) 
  ) 
  
  X_ext$x4 = round(X_ext$x4) + 1
  X_ext$x5 = 50 + 10 * X_ext$x1 + (2) * X_ext$x2 + (-1) * X_ext$x3 + (-0.3) * X_ext$x4 + rnorm(100, mean = 0, sd = 10)
  
  varnames = c("1", paste0("x", 1:5))
  
  #============ Specify outcome models ==============
  model_form_x_t1 = setNames(c(10.0, 0.05, -1.5, -1.0, -0.2, -0.1), varnames) # 1.5*A, sigma = 4.0
  model_form_x_t2 = setNames(c(6.0, 0.5, -0.5, -1.0, -0.3, -0.06), varnames) # 1.8*A, sigma = 4.0
  model_form_x_t3 = setNames(c(5.0, 1.9, 1.4, -1.3, -0.4, -0.15), varnames) # 1.6*A, sigma = 4.0
  model_form_x_t4 = setNames(c(1.2, 1.0, 2.0, -0.5, -0.4, -0.10), varnames) # 2.5*A, sigma = 5.0
  
  outcome_model_specs = list(
    list(effect = 0, model_form_x = model_form_x_t1, # from data: true_effect = 1.5
         noise_mean = 0, noise_sd = 4),  # model form for the first time point, given by model_form_x_t1
    list(effect = 0, model_form_x = model_form_x_t2, # from data: true_effect = 1.8
         noise_mean = 0, noise_sd = 4),  # model form for the second time point, given by model_form_x_t2
    list(effect = 0, model_form_x = model_form_x_t3, # from data: true_effect = 1.6
         noise_mean = 0, noise_sd = 4),  # model form for the third time point, given by model_form_x_t3
    list(effect = true_effect_long, model_form_x = model_form_x_t4, # from data: true_effect = 2.5
         noise_mean = 0, noise_sd = 4)  # model form for the fourth time point, given by model_form_x_t4
  )
  
  #=========== generate trial data ============
  Data = simulate_trial(X_int, 
                        X_ext, 
                        num_treated = 150, 
                        OLE_flag = T, 
                        T_cross = 2,  
                        outcome_model_specs) 
  

  data_matrix_list[[trial_iter]] = Data
}
```




### 2 Bootstrap inference
```{r}
# Specify a list of methods to be tested
bootstrap_obj = setup_bootstrap(
  replicates = 1e3,
  bootstrap_CI_type = "perc"
)

model_form_mu = c("y1 ~ x1 + x2 + x3 + x4 + x5", 
                  "y2 ~ x1 + x2 + x3 + x4 + x5",
                  "y3 ~ x1 + x2 + x3 + x4 + x5",
                  "y4 ~ x1 + x2 + x3 + x4 + x5")

## IPW - optimal weight
method_IPW_DID = setup_method_DID(method_name = "IPW",
                                  bootstrap_flag = T,
                                  bootstrap_obj = bootstrap_obj,
                                  model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
                                  model_form_piA = "A ~ x1 + x2 + x3 + x4 + x5")

## AIPW - optimal weight
method_AIPW_DID = setup_method_DID(method_name = "AIPW",
                                   bootstrap_flag = T,
                                   bootstrap_obj = bootstrap_obj,
                                   model_form_piS = "S ~ x1 + x2 + x3 + x4 + x5",
                                   model_form_piA = "A ~ x1 + x2 + x3 + x4 + x5",
                                   model_form_mu0_ext = model_form_mu)

## OR - optimal weight

method_OR_DID = setup_method_DID(method_name = "OR",
                                 bootstrap_flag = T,
                                 bootstrap_obj = bootstrap_obj,
                                 model_form_mu0_ext = model_form_mu,
                                 model_form_mu0_rct = model_form_mu,
                                 model_form_mu1_rct = model_form_mu)

## method list
method_obj_list = list(method_IPW_DID, 
                       method_AIPW_DID,
                       method_OR_DID)

```



```{r}
# create a simulation object for primary analysis
simulation_OLE_obj = setup_simulation_OLE(
  data_matrix_list = data_matrix_list,  # two scenarios
  trial_status_col_name = trial_status_col_name, 
  treatment_col_name = treatment_col_name, 
  outcome_col_name = outcome_col_name, 
  covariates_col_name = covariates_col_name, 
  method_obj_list = method_obj_list, 
  true_effect = true_effect_long,  
  T_cross = 2,
  alpha = alpha,
  method_description = c("IPW, DID", 
                         "AIPW, DID", 
                         "OR, DID"))
```

```{r}
simulation_report = run_simulation(simulation_OLE_obj, quiet = F)
```

```{r}
simulation_report # Type I error and Power
# simulation hypo test
```